Increased perceptions of autonomy through choice fail to enhance motor skill retention

### 2021-11-03 docs/r1_submission/r1_responses-to-reviewers.Rmd

output      : papaja::revision_letter_pdf
---

Jacqueline Clare Snow, PhD  
Associate Editor  
*Journal of Experimental Psychology: Human Perception and Performance*

Dear Dr. Snow,

Thank you for taking the time to consider our manuscript for publication at _`r rmarkdown::metadata$journal`_. We have read the comments provided by the Reviewers and have revised the manuscript accordingly. We thank the Reviewers for their helpful and insightful comments. Below we provide a point-by-point reply to the Reviewer comments (in bold-italic) and our responses in normal text. Any substantial changes that have been made to the original submission are also included in the response letter (in a text box) and are in red font in the revised manuscript.

@@

# Reviewer \#2

\RC{\emph{The authors submitted a well written manuscript reporting a pre-registered experiment that investigated the effect of awareness of being denied control over two relevant variables of the practice environment (i.e., frequency and speed of video demonstration) as participants learned the speed cup-stacking task. Results revealed that participants allowed to choose when to watch and the speed of the video demonstrations (i.e., self-controlled group) showed higher perceived autonomy scores when compared to yoked participants who received the demonstrations in matched schedule to a self-controlled counterpart and were either explicitly aware of this information (Explicit Yoked) or not (Traditional Yoked). However, the authors failed to replicate the so-called self-control learning benefit as no differences between groups were found in the retention test. Additionally, groups did not differ in self-reported measures of perceived competence and intrinsic motivation.}

\emph{Overall, I think this paper can be a valuable addition to the literature. The study is methodologically strong, it has one of the largest sample sizes reported in the motor learning literature, the authors pre-registered the main analyses, carried out a power calculation, and equivalence test, and heartfully embraced open science practices (I was able to easily reproduce the analyses/figures reported in the paper). However, I do have a few concerns/recommendations that, if addressed, can help to further improve the paper. 
}}

Thank your for the positive general comments about our manuscript. We are pleased to hear that you were able to reproduce the analyses and figures. Below are the responses to your specific comments. We have split the first comment into smaller chunks to facilitate our response.

@@

\RC{\emph{My major concern is how the study's rationale and relevance are framed. The authors start by questioning the role played by autonomy support in practice conditions wherein provision of choice is manipulated. They argued that "if the benefits are the result of having opportunities for choice then learning differences should not emerge between different self-controlled groups" [Page 5, line 7]. However, this claim disregards another possible explanation for the self-control learning benefit, namely information processing factors (Barros et al., 2019).
}}

We agree with the reviewer that an information-processing explanation has also been forwarded to account for the self-controlled learning advantages. We did not not specifically refer to it in the introduction as this experiment was not designed to dissociate between the motivational and the information-processing explanations. We have removed the quoted text in the revised submission and have added the following as a footnote to acknowledge the information-processing view **(see Page XX)**:

> It should be noted that other researchers have instead presented an information-processing explanation for the self-controlled learning advantages (see Ste-Marie et al., 2019 for a recent discussion of the motivational and information-processing explanations). We acknowledge this view here; however, unlike previous experiments (e.g., Barros, Yantha, Carter, Hussien, & Ste-Marie, 2019; Carter, Carlsen, & Ste-Marie, 2014; Carter & Ste-Marie, 2017b, 2017a; Couvillion, Bass, & Fairbrother, 2020; Woodard & Fairbrother, 2020) the current experiment was not designed to test between explanations. We focused on the motivational explanation as this view has garnered more attention due the “OPTIMAL” theory of motor learning.

\RC{\emph{Second, they state that "there has been little-to-no support for the notion that practicing in a self-controlled group is perceived as more autonomy-supportive than being in a yoked group" [Page 5, line 15]. Although this statement has been supported by previous studies such as the one cited in the paper (Ste-Marie et al., 2013), group differences regarding levels of perceived autonomy have been found in recent studies (McKay \& Ste-Marie, 2020- also cited in the paper). The evidence, therefore, as presented in the paper seems mixed. Moreover, another explanation for the lack of group differences might have to do with the fact that increased levels of perceived autonomy have been assumed instead of directly measured in self-control studies. Thus, it becomes unclear whether the little-to-no support for the notion that practicing in a self-controlled group is perceived as more autonomy-supportive comes from a lack of differences when there is a measure of perceived autonomy or, rather, from a lack of studies assessing how provision of choice affects perception of autonomy. The authors then proceed to claim that "Although the available literature does not provide substantial evidence that self-controlled practice is autonomy-supportive, this does not necessarily mean such [extra 'a' here] manipulation is not autonomy-supportive. For instance, this effect may in fact be quite small and require much larger sample sizes to detect than those commonly used in motor learning experiments" [Page 5, line 25], which is a fair claim to make considering the methodological concerns surrounding motor learning studies (Lohse et al., 2016). However, the way this argument was structured led to the notion that it is still unclear whether self-control conditions lead to higher levels of perceived autonomy, circling back to my previous argument that the evidence is somewhat mixed and might be due to the lack of studies directly measuring perceived autonomy.
}}

We agree with many of the points raised here. We attempted to capture the mixed nature of findings by presenting experiments that failed to find an effect and the one that did find the expected effect. We restricted our discussion of past work not including measures in the Discussion as the true impact of this unknown and speculative. Based on your comments, we have restructured part of the Introduction to improve the clarity of our arguments **(see Pages XX, Lines XX)**:

> Despite its prominent role as a robust and generalizable learning variable in the “OPTIMAL” theory (p. 1393), there is considerable ambiguity surrounding whether the provision of choice is in fact an autonomy-supportive manipulation. First, the notion that practicing in a self-controlled group is actually more autonomy-supportive than a yoked group has primarily been assumed (e.g., Abdollahipour, Palomo Nieto, Psotta, & Wulf, 2017; Chua, Wulf, & Lewthwaite, 2018; Lewthwaite, Chiviacowsky, Drews, & Wulf, 2015; Wulf, Lewthwaite, Cardozo, & Chiviacowsky, 2018) rather than supported empirically. Second, when researchers have included measures related to perceptions of autonomy the data is mixed. For example, Ste-Marie, Vertes, Law, and Rymal (2013) did not find the expected effect of higher perceptions of autonomy during practice in a self-controlled group as compared to a yoked group. Similar outcomes have been reported by others (e.g., Barros et al., 2019; Carter & Ste-Marie, 2017b; McKay & Ste-Marie, 2020a). In contrast, McKay and Ste-Marie (2020b) recently found that practicing in a self-controlled group was perceived as more autonomy-supportive than practicing the same task in a yoked group. However, the higher perceived autonomy scores did not translate into enhanced motor performance or learning in the self-controlled group as predicted in the “OPTIMAL” theory. Although the majority of experiments that included a measure related to perceived autonomy reported no group differences, the “absence of evidence is not evidence of absence” (Altman & Bland, 1995). Thus, self-controlled practice conditions could be an autonomy-supportive manipulation but such an effect may actually be quite small and require much larger sample sizes to detect than those used in previous experiments (e.g., Barros et al., 2019; Ste-Marie et al., 2013) and in motor learning experiments in general (see Lohse, Buchanan, & Miller, 2016 for a discussion). This argument of underpowered experimental designs is supported by the results of McKay and Ste-Marie (2020b) as these authors had one of the largest sample sizes to date in the self-controlled literature.

\RC{\emph{Moreover, there was no explicit mention as to how levels of perceived autonomy might have a mechanistic role in motor skill learning. Self-control conditions might lead to higher levels of perceived autonomy but no learning advantage (as seen in the present study). But, when looking at the title of the paper and how the discussion unfolds, the authors shift their focus to claim that autonomy-support does not benefit motor learning.
}}

We agree that self-controlled practice conditions might lead to higher levels of perceived autonomy without any learning advantage. This, however, is not what is predicted in "OPTIMAL" theory where a causal relationship is put forward. The title of our paper and how the discussion unfolds is based on the results of our experiment, which suggests that increased perceptions of autonomy do not benefit motor learning. To your first point, we have added the following to the introduction to better outline how autonomy-support might have a mechanistic role in motor performance and learning according to the "OPTIMAL" theory of motor learning **(see Pages XX, Lines XX)**:

> Within their “OPTIMAL” theory of motor learning, Wulf and Lewthwaite (2016) argued that providing learners with opportunities for choice creates an autonomy-supportive practice environment, which facilitates motor performance and learning. Specifically, the authors predict that autonomy-support facilitates performance by enhancing expectancies (Prediction 3, p. 1404), that enhanced expectancies and autonomy support contribute to efficient goal-action coupling by readying the motor system for task execution (Prediction 2, p. 1404), and that enhanced expectancies and autonomy support facilitate motor learning by making dopamine available for memory consolidation and neural pathway development (Prediction 7, p. 1404). In other words, these psychological benefits of increased perceptions of autonomy and competence, and the resulting increases in performance and learning are a by-product of having choice itself. Overall, Wulf and Lewthwaite (2016)’s “OPTIMAL” theory of motor learning provides a motivational explanation for the learning advantages of self-controlled practice conditions over yoked practice conditions.

\RC{\emph{The narrative in its current form lacks clarity. Are the authors interested in detecting group differences in levels of perceived autonomy (regardless of whether it predicts learning) or in the role of autonomy-support in enhancing learning (i.e., replicate the self-control learning benefit under a motivational perspective)? The issue gets even more cloudy as the authors raise two methodological concerns identified in the autonomy-support literature: 1) "The first of these is that in self-controlled motor learning experiments participants in the self-controlled group are usually given choice over a single component of their practice and participants in the yoked group are not given choice over this component. However, within the context of practice itself there are many other opportunities for choice that participants may explore, independent of their assigned group" [Page6, line 4], and 2) "The other consideration relates to the instructions that are provided to participants in the yoked group [Page 6, line 16]", "Participants in yoked groups are not even aware that they have been denied an opportunity for choice, nor that their feedback schedule was created by another participant who was given choice over when feedback was or was not provided" [Page 6, line 20]. Regarding the first one, the benefits of self-control experiments have been found even when choice is given over an irrelevant aspect of the practice condition (incidental choice - Wulf \& Lewthwaite, 2016). Following this logic, it seems like the authors are referring specifically to the levels of perceived autonomy (having control over more variables might lead to higher levels of perceived autonomy). However, making yoked participants aware of being denied choice over some aspects of the practice condition does not prevent them from experiencing choice through other opportunities, weaking this argument. Moreover, this manipulation (i.e., making yoked participants aware of being denied choice) is better fitted under the controlling instructional language manipulation, which is contrasted against autonomy-supportive language (e.g., Hooyman et al., 2014). Yet, the authors do not even mention past research investigating the effect of controlling language on motor learning.
}}

Based on the "OPTIMAL" theory, we were interested in replicating the typical self-controlled learning advantage (group difference in retention favouring the self-controlled group) and testing the claim that self-controlled practice conditions are a way to create an autonomy-supportive environment (group difference in self-reported perceptions of autonomy). We have added details about our predictions to the revised manuscript (but we include this information below with your specific comment about these missing from the manuscript). In terms of the first methodological limitation, yes some researchers have reported a learning benefit of task-irrelevant choices (e.g., Lewthwaite & Wulf 2015); however, this is a fairly selective stance given the mixed results surrounding task-irrelevant choices (e.g., Carter & Ste-Marie 2017; Grand et al. 2015; McKay & Ste-Marie 2020ab). In McKay & Ste-Marie 2020a, the task-irrelevant choice groups actually performed with significantly more error in the acquisition and transfer.

### 2021-11-04 docs/r1_submission/r1_manuscript.Rmd
update: finalized changes to revised manuscript

  ***Public Significance Statement***  
  The present study suggests that although university-aged students may experience a boost in feelings of autonomy when given choice over features of their learning environment, this boost does not enhance learning of a new motor skill. Additionally, it highlights that coaches and practictioners should be encouraged to incorporate other and more effective practice variables into their training regimes.

keywords          : "Motor learning; Pre-registered; Self-controlled; Observation; Equivalence testing"
wordcount         : "5055"
bibliography      : ["../references.bib", "../r-references.bib"]

@@

Over the years, this manipulation has been described using a variety of names (e.g., self- and/or learner-controlled; -regulated; -directed; -selected), but more recently it has been subsumed by autonomy-support. \textcolor{red}{Within their ``OPTIMAL'' theory of motor learning,} @wulf2016 \textcolor{red}{argued that providing learners with opportunities for choice creates an autonomy-supportive practice environment, which facilitates motor performance and learning. Specifically, the authors predict that autonomy-support facilitates performance by enhancing expectancies (Prediction 3, p. 1404), that enhanced expectancies and autonomy support contribute to efficient goal-action coupling by readying the motor system for task execution (Prediction 2, p. 1404), and that enhanced expectancies and autonomy support facilitate motor learning by making dopamine available for memory consolidation and neural pathway development (Prediction 7, p. 1404). In other words, these psychological benefits of increased perceptions of autonomy and competence, and the resulting increases in performance and learning are a by-product of having choice itself. Overall,} @wulf2016's \textcolor{red}{``OPTIMAL'' theory of motor learning provides a motivational explanation for the learning advantages of self-controlled practice conditions over yoked practice conditions.}[^1]

@@

\textcolor{red}{Despite its prominent role as a robust and generalizable learning variable in the ``OPTIMAL'' theory (p. 1393), there is considerable ambiguity surrounding whether the provision of choice is in fact an autonomy-supportive manipulation. First, the notion that practicing in a self-controlled group is actually more autonomy-supportive than a yoked group has primarily been assumed} [e.g., @abdollahipour2017; @chua2018; @lewthwaite2015; @wulf2018] \textcolor{red}{rather than supported empirically. Second, when researchers have included measures related to perceptions of autonomy the data is mixed. For example,} @ste-marie2013 \textcolor{red}{did not find the expected effect of higher perceptions of autonomy during practice in a self-controlled group as compared to a yoked group. Similar outcomes have been reported by others} [e.g., @barros2019; @carter2017a; @mckay2020b]. \textcolor{red}{In contrast,} @mckay2020a \textcolor{red}{recently found that practicing in a self-controlled group was perceived as more autonomy-supportive than practicing the same task in a yoked group. However, the higher perceived autonomy scores did not translate into enhanced motor performance or learning in the self-controlled group as predicted in the ``OPTIMAL'' theory. Although the majority of experiments that included a measure related to perceived autonomy reported no group differences, the ``absence of evidence is not evidence of absence''} [@altman1995]\textcolor{red}{. Thus, self-controlled practice conditions could be an autonomy-supportive manipulation but such an effect may actually be quite small and require much larger sample sizes to detect than those used in previous experiments} [e.g., @barros2019; @ste-marie2013] \textcolor{red}{and in motor learning experiments in general} [see @lohse2016 for a discussion]\textcolor{red}{. This argument of underpowered experimental designs is supported by the results of} @mckay2020a \textcolor{red}{as these authors had one of the largest sample sizes to date in the self-controlled literature.}

@@

Here we investigated the effects of making participants in a yoked group explicitly aware of not only being denied opportunities for choice over their frequency of watching and the playback speed of video demonstrations, but also that the schedule they would experience during practice was created by another participant in the experiment. The addition of this novel yoked group allowed us to address one of the previously identified methodological limitations regarding experimental group instructions in past self-controlled research. We compared the performance of this explicit yoked group with traditional self-controlled and yoked groups on a speed cup-stacking task in practice and in a delayed retention test.

@@

One-hundred and fifty participants completed the experiment. Sample size was determined by an a priori power calculation based on our smallest comparison of interest. An early estimate for the effect of self-controlled over yoked practice was a Hedges' $g = 0.63$ [@mckay2014] and while planning this experiment this effect was estimated to be $g = 0.52$ (Z. Yantha, personal communication, October 2019). Based on this effect, a positive correlation of $r = 0.6$ between retention and pre-test as the covariate, and 80% power to detect a difference between the self-controlled and traditional yoked groups, the required number of participants was 31 per group. Considering the novelty of our Explicit Yoked group, we assumed a smaller effect, $g = 0.4$, between it and the Traditional Yoked group. With the same parameters as above, this resulted in our final sample of 50 participants per group.

Participants completed the experiment in either the Self-Controlled group $(M_{age} = 18.0$, *SD* = 0.34), the Traditional Yoked group $(M_{age} = 19.5$, *SD* = 1.89), or the Explicit Yoked group $(M_{age} = 19.2$, *SD* = 1.55). We collected the Self-Controlled group first as their self-selected observation schedule was required for the yoking procedure for the two other groups. Once the Self-Controlled group had been collected, participants were randomly assigned to one of the two Yoked groups. All participants provided written informed consent approved by and conducted in accordance with the University's Research Ethics Board. Participants received either \$15 or a course bonus for their participation.

@@

To test predictions based on the "OPTIMAL" theory regarding the role of motivation, enhanced expectancies, and autonomy-support, participants completed the interest/enjoyment and perceived competence subscales from the Intrinsic Motivation Inventory [@mcauley1989] and a custom scale regarding choice used in previous self-controlled motor learning experiments [@barros2019; @carter2017a]. The order of questions from each scale were randomized and each question was rated using a 7-point Likert scale. Participants answered these questions after the pre-test, after trials five and 25 of acquisition, and before the delayed retention test.

@@

@@

Our primary outcome measure was stacking time (i.e., response time) in seconds. Trials recorded as errors $(76/5250 = 1.45\%)$ during data collection were manually removed prior to data analysis. For each participant, pre-test and delayed retention trials were aggregated into one block of five trials and acquisition was aggregated into five blocks of five trials. Significance level was set to 0.05 for all statistical tests. Effect sizes for omnibus tests are reported using generalized eta squared $(\eta^2_{G})$ or eta squared $(\eta^2)$. Post hoc comparisons were conducted using Holm's correction. A Cook's distance of $\geq 1$ was used to identify any influential cases and none were identified.

@@

To test whether delayed retention (Figure \@ref(fig:fig1)C) was differentially impacted by the experimental group (i.e., Self-Controlled, Traditional Yoked, Explicit Yoked) experienced during acquisition, we performed a one-way ANCOVA controlling for pre-test. As can be seen, the Self-Controlled group $(M = 9.99$, $95\%CI = [9.68,10.31])$, the Traditional Yoked group $(M = 10.18$, $95\%CI = [9.86,10.50])$, and the Explicit Yoked group $(M = 10.12$, $95\%CI = [9.81,10.44])$ all had similar stacking times in retention (means are shown as the adjusted means controlling for pre-test). The effect of Group was not significant, $F(2,146) = .335$, $p = .716$, $\eta^2 = .002$.

@@

table_2 <- tibble(
    Observation = c("Model viewing frequency",
                    "Regular speed frequency"),
    b1 = c("32.6%",
           "33.8%"),
    b2 = c("16.9%",
           "38.1%"),
    b3 = c("12.0%%",
           "50.0%"),
    b4 = c("12.9%",
           "34.4%"),
    b5 = c("9.3%",
           "30.4%"),
)

kbl(table_2,
    booktabs = TRUE,
    escape = TRUE,
    linesep = "\\addlinespace",
    caption = "The mean frequencies for each Acquisition block of requesting to watch the video demonstration and of those requests, the mean frequency of choosing the regular viewing speed.",
    col.names = c(" ",
                  "Block 1",
                  "Block 2",
                  "Block 3",
                  "Block 4",
                  "Block 5")
    ) %>% 
    kable_styling(position = "left",
                  font_size = 11) %>% 
    column_spec(1, width = "13em") %>% 
    column_spec(2, width = "5em") %>% 
    column_spec(3, width = "5em") %>% 
    column_spec(4, width = "5em") %>% 
    column_spec(5, width = "5em")

@@

Model frequency and the video speed request data are reported in Table \@ref(tab:table2). The mean frequency of viewing the video demonstration during acquisition was $16.7\%$. When a video demonstration was selected, the mean frequency that the regular speed was selected was $37.3\%$. Stacking time for each Group decreased across acquisition trials (Figure \@ref(fig:fig1)A) and blocks (Figure \@ref(fig:fig1)B). A 3 Group x 5 Block mixed ANOVA with repeated measures on Block revealed a significant Block effect, $F(3.63,533.79) = 116.73$, $p < .001$, $\eta^2_{G} = .133$. Post hoc comparisons revealed all acquisition blocks were significantly different from each other. Both the Group effect, $F(2,147) = .738$, $p = .48$, $\eta^2_{G} = .008$, and Group x Block interaction, $F(7.26,533.79) = .543$, $p = .802$, $\eta^2_{G} = .001$, were not statistically significant.

@@

To further probe the null effects [@harms2018] related to our main predictions, two equivalence tests (Self-Controlled versus Traditional Yoked and Traditional Yoked versus Explicit Yoked) on pre-test adjusted retention stacking times were carried out using the two one-sided test procedure [@lakens2018; @schuirmann1987]. To establish our smallest effect size of interest [@lakens2021] we calculated the effect size our design had 33% power to detect [@simonsohn2015], which was $d_{s} = .31$. The Self-Controlled and Traditional Yoked groups were not statistically different, $t(97.99) = -0.830$, $p = 0.409$, and not statistically equivalent, $t(97.99) = 0.720$, $p = 0.237$, given equivalence bounds of $-.355$ and $.355$ (on a raw scale). Similarly, the Traditional Yoked and Explicit Yoked groups were not statistically different, $t(97.99) = 0.262$, $p = 0.794$, and not statistically equivalent, $t(97.99) = -1.288$, $p = 0.100$, given equivalence bounds of $-.355$ and $.355$ (on a raw scale).

>>>>

  \noindent \textbf{\emph{Public Significance Statement}}  
  The present study suggests that although university-aged students may experience a boost in feelings of autonomy when given choice over features of their learning environment, this boost may not enhance learning of a new motor skill. Our results also highlight the need for higher powered and pre-registered motor learning experiments if coaches and practictioners are to be provided with reliable recommendations for training and rehabilitation protocols.

keywords          : "Motor learning; Pre-registered; Self-controlled; Observation"
wordcount         : "4585"
bibliography      : ["../references.bib", "../r-references.bib"]

@@

Over the years, this manipulation has been described using a variety of names (e.g., self- and/or learner-controlled; -regulated; -directed; -selected), but more recently \textcolor{red}{some researchers have adopted the term} autonomy-support. \textcolor{red}{Within their ``OPTIMAL'' theory of motor learning,} @wulf2016 \textcolor{red}{argued that providing learners with opportunities for choice creates an autonomy-supportive practice environment, which facilitates motor performance and learning. Specifically, the authors predict that autonomy-support facilitates performance by enhancing expectancies (Prediction 3, p. 1404), that enhanced expectancies and autonomy support contribute to efficient goal-action coupling by readying the motor system for task execution (Prediction 2, p. 1404), and that enhanced expectancies and autonomy support facilitate motor learning by making dopamine available for memory consolidation and neural pathway development (Prediction 7, p. 1404). In other words, these psychological benefits of increased perceptions of autonomy and competence, and the resulting increases in performance and learning are a by-product of having choice itself. Overall,} @wulf2016's \textcolor{red}{``OPTIMAL'' theory of motor learning provides a motivational explanation for the learning advantages of self-controlled practice conditions over yoked practice conditions.}[^1]

@@

\textcolor{red}{Despite its prominent role as a robust and generalizable learning variable in the ``OPTIMAL'' theory (p. 1393), there is considerable ambiguity surrounding whether the provision of choice is in fact an autonomy-supportive manipulation. First, the notion that practicing in a self-controlled group is actually more autonomy-supportive than a yoked group has primarily been assumed} [e.g., @abdollahipour2017; @chua2018; @lewthwaite2015; @wulf2018] \textcolor{red}{rather than supported empirically. Second, when researchers have included measures related to perceptions of autonomy the data is mixed. For example,} @ste-marie2013 \textcolor{red}{did not find the expected effect of higher perceptions of autonomy during practice in a self-controlled group as compared to a yoked group. Similar outcomes have been reported by others} [e.g., @barros2019; @carter2017a; @mckay2020b]. \textcolor{red}{In contrast,} @mckay2020a \textcolor{red}{recently found that practicing in a self-controlled group was perceived as more autonomy-supportive than practicing the same task in a yoked group. However, despite their higher perceived autonomy scores the self-controlled group did not have significantly better motor performance and learning as predicted in the ``OPTIMAL'' theory. Although the majority of experiments that included a measure related to perceived autonomy reported no group differences, the ``absence of evidence is not evidence of absence''} [@altman1995]\textcolor{red}{. Thus, self-controlled practice conditions could be an autonomy-supportive manipulation but such an effect may actually be quite small and require much larger sample sizes to detect than those used in previous experiments} [e.g., @barros2019; @ste-marie2013] \textcolor{red}{and in motor learning experiments in general} [see @lohse2016 for a discussion]\textcolor{red}{. This argument of underpowered experimental designs is supported by the results of} @mckay2020a \textcolor{red}{as these authors had one of the largest sample sizes to date in the self-controlled literature.}

@@

\textcolor{red}{Here, we addressed the methodological limitation of past self-controlled research where participants in the yoked group are unaware that a feature of their practice environment was created by another participant in the experiment. To this end, participants learned a speed cup-stacking task in either a self-controlled group, a traditional yoked group, or a novel explicitly aware yoked group. The self-controlled group has choice over the frequency of watching a video demonstration and the video playback speed (real-time or slow motion). Participants in the traditional and explicit yoked groups were matched to a participant in the self-controlled group and experienced the observation schedule selected by this participant. The key difference being that participants in the explicit yoked group were told that the observation schedule they would experience during practice was selected by another participant, whereas participants in the traditional yoked group were not aware of this. Motor learning was assessed using a delayed retention test. We predicted that the self-controlled group would have significantly faster stacking times in retention than the traditional yoked group. In other words, we expected to replicate the typical self-controlled learning advantage. If the self-controlled learning advantage results from choice being autonomy-supportive as argued in the ``OPTIMAL'' theory} [@wulf2016]\textcolor{red}{, then participants in the self-controlled group should also self-report significantly higher scores for perceptions of autonomy. Also based on the view that autonomy-support is the mechanism underlying the typical self-controlled learning advantage} [@wulf2016]\textcolor{red}{, we predicted that the explicit yoked group should have significantly slower stacking times in retention. Although we found significantly higher perceptions of autonomy in the self-controlled group, there were no significant group differences in stacking times during retention. Overall, these results do not support our predictions and are inconsistent with predictions 2, 3, and 7 in the ``OPTIMAL'' theory of motor learning} [@wulf2016].

@@

One-hundred and fifty \textcolor{red}{university students} participated in the experiment. Sample size was determined by an a priori power calculation based on our smallest comparison of interest \textcolor{red}{using GLIMMPSE (https://glimmpse.samplesizeshop.org/)}. An early estimate for the effect of self-controlled over yoked practice was a Hedges' $g = 0.63$ [@mckay2014] and while planning this experiment this effect was estimated to be $g = 0.52$ (Z. Yantha, personal communication, October 2019). Based on this effect, a positive correlation of $r = 0.6$ between retention and pre-test as the covariate, \textcolor{red}{an alpha of 0.05}, and 80% power to detect a difference between the self-controlled and traditional yoked groups, the required number of participants was 31 per group. Considering the novelty of our Explicit Yoked group, we assumed a smaller effect, $g = 0.4$, between it and the Traditional Yoked group. \textcolor{red}{Using this effect and} the same parameters as above, this resulted in our final sample of 50 participants per group.

Participants completed the experiment in either the Self-Controlled group $(M_{age} = 18.0$, *SD* = 0.34; \textcolor{red}{32 females}), the Traditional Yoked group $(M_{age} = 19.5$, *SD* = 1.89; \textcolor{red}{28 females}), or the Explicit Yoked group $(M_{age} = 19.2$, *SD* = 1.55; \textcolor{red}{30 females}). We collected the Self-Controlled group first as their self-selected observation schedule was required for the yoking procedure for the two other groups. Once the Self-Controlled group had been collected, participants were randomly assigned to one of the two Yoked groups. All participants provided written informed consent approved by and conducted in accordance with the University's Research Ethics Board. Participants received either \$15 or a course bonus for their participation.

@@

To test predictions based on the "OPTIMAL" theory regarding the role of motivation, enhanced expectancies, and autonomy-support, participants completed the interest/enjoyment and perceived competence subscales from the Intrinsic Motivation Inventory [@mcauley1989] and a custom scale regarding choice used in previous self-controlled motor learning experiments [@barros2019; @carter2017a]. The order of questions from each scale were randomized and each question was rated using a 7-point Likert scale. Participants answered these questions after the pre-test, after trials five and 25 of acquisition, and before the delayed retention test. The values for Cronbach's alpha for each questionnaire at each time point are reported in Table \@ref(tab:table2).

@@

table_2 <- tibble(
    c1 = c("Perceived autonomy",
           "Perceived competence",
           "Intrinsic motivation"),
    c2 = c("0.62",
           "0.90",
           "0.86"),
    c3 = c("0.79",
           "0.92",
           "0.90"),
    c4 = c("0.79",
           "0.94",
           "0.92"),
    c5 = c("0.84",
           "0.92",
           "0.92"),
)

kbl(table_2,
    booktabs = TRUE,
    escape = TRUE,
    linesep = "\\addlinespace",
    caption = "Cronbach's alpha for each questionnaire at each time point.",
    col.names = c(" ",
                  "After pre-test",
                  "After trial 5",
                  "After trial 25",
                  "Before retention")
    ) %>% 
    kable_styling(position = "left",
                  font_size = 11) %>% 
    column_spec(1, width = "10em") %>% 
    column_spec(2, width = "6em") %>% 
    column_spec(3, width = "6em") %>% 
    column_spec(4, width = "6em") %>% 
    column_spec(5, width = "7em")

@@

Our primary outcome measure was stacking time (i.e., response time) in seconds. Trials recorded as errors $(76/5250 = 1.45\%)$ during data collection were manually removed prior to data analysis. For each participant, pre-test and delayed retention trials were aggregated into one block of five trials and acquisition was aggregated into five blocks of five trials. Significance level was set to 0.05 for all statistical tests. Effect sizes for omnibus tests are reported using generalized eta squared $(\eta^2_{G})$ or eta squared $(\eta^2)$. Post hoc comparisons were conducted using Holm's correction. A Cook's distance of $\geq 1$ was used to identify any influential cases and none were identified. \textcolor{red}{Statistical tests are described below.}

@@

To test whether delayed retention (Figure \@ref(fig:fig1)C) was differentially impacted by the experimental group experienced during acquisition, we performed a one-way ANCOVA controlling for pre-test. As can be seen, the Self-Controlled group $(M = 9.99$, $95\%CI = [9.68,10.31])$, the Traditional Yoked group $(M = 10.18$, $95\%CI = [9.86,10.50])$, and the Explicit Yoked group $(M = 10.12$, $95\%CI = [9.81,10.44])$ all had similar stacking times in retention (means are shown as the adjusted means controlling for pre-test). The effect of Group was not significant, $F(2,146) = .335$, $p = .716$, $\eta^2 = .002$.

@@

```{r table3, echo = FALSE, results = "asis"}

table_3 <- tibble(
    Observation = c("Model viewing frequency",
                    "Regular speed frequency"),
    b1 = c("32.6%",
           "33.8%"),
    b2 = c("16.9%",
           "38.1%"),
    b3 = c("12.0%%",
           "50.0%"),
    b4 = c("12.9%",
           "34.4%"),
    b5 = c("9.3%",
           "30.4%"),
)

kbl(table_3,
    booktabs = TRUE,
    escape = TRUE,
    linesep = "\\addlinespace",
    caption = "The mean frequencies for each Acquisition block of requesting to watch the video demonstration and of those requests, the mean frequency of choosing the regular viewing speed.",
    col.names = c(" ",
                  "Block 1",
                  "Block 2",
                  "Block 3",
                  "Block 4",
                  "Block 5")
    ) %>% 
    kable_styling(position = "left",
                  font_size = 11) %>% 
    column_spec(1, width = "13em") %>% 
    column_spec(2, width = "5em") %>% 
    column_spec(3, width = "5em") %>% 
    column_spec(4, width = "5em") %>% 
    column_spec(5, width = "5em")

@@

Model frequency and the video speed request data are reported in Table \@ref(tab:table3). The mean frequency of viewing the video demonstration during acquisition was $16.7\%$. When a video demonstration was selected, the mean frequency that the regular speed was selected was $37.3\%$. Stacking time for each Group decreased across acquisition trials (Figure \@ref(fig:fig1)A) and blocks (Figure \@ref(fig:fig1)B). A 3 Group x 5 Block mixed ANOVA with repeated measures on Block revealed a significant Block effect, $F(3.63,533.79) = 116.73$, $p < .001$, $\eta^2_{G} = .133$. Post hoc comparisons revealed all acquisition blocks were significantly different from each other. Both the Group effect, $F(2,147) = .738$, $p = .48$, $\eta^2_{G} = .008$, and Group x Block interaction, $F(7.26,533.79) = .543$, $p = .802$, $\eta^2_{G} = .001$, were not statistically significant.

@@

\textcolor{red}{Given our pre-registered analysis resulted in a non-significant finding, we performed two equivalence tests (Self-Controlled versus Traditional Yoked and Traditional Yoked versus Explicit Yoked) on pre-test adjusted retention stacking times using the two one-sided test procedure} [@lakens2018; @schuirmann1987]. \textcolor{red}{Equivalence tests are a statistical tool that researchers can use to support the absence of a meaningful effect and avoid incorrectly concluded no effect exists based on a non-significant finding} [@harms2018]. To establish our smallest effect size of interest [@lakens2021] we calculated the effect size our design had 33% power to detect [@simonsohn2015], which was $d_{s} = .31$. The Self-Controlled and Traditional Yoked groups were not statistically different, $t(97.99) = -0.830$, $p = 0.409$, and not statistically equivalent, $t(97.99) = 0.720$, $p = 0.237$, given equivalence bounds of $-.355$ and $.355$ (on a raw scale). Similarly, the Traditional Yoked and Explicit Yoked groups were not statistically different, $t(97.99) = 0.262$, $p = 0.794$, and not statistically equivalent, $t(97.99) = -1.288$, $p = 0.100$, given equivalence bounds of $-.355$ and $.355$ (on a raw scale).
